---
title: "CCMnet: Edge, Degree, Mixing, and Structural Assessment"
author: "Ravi Goyal"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCMnet Combined Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4
)

# Load libraries
library(CCMnet)
library(tidyverse)
library(igraph)
library(gtools)
library(mvtnorm)
```

# Section 1: Introduction

This vignette demonstrates we test whether a network property (e.g., degree distribution) is part of the generative process by comparing observed networks to Null CCMs.

# Section 2: Structural Assessment

Structural assessment tests whether a network property is part of the generative process.

- **Erdős-Rényi (ER) Network:** Only the number of edges is relevant; degree distribution deviations should be minimal.
- **Barabási–Albert (BA) Network:** Preferential attachment induces heterogeneous degrees; a Null model with edges only cannot capture this.

## Example 2.1: ER Network

```{r structural-er}
population <- 100
p_edge <- 0.05
G_obs <- erdos.renyi.game(n = population, p.or.m = p_edge, type = "gnp", directed = FALSE)

# Number of edges and empirical probability
num_edges_obs <- gsize(G_obs)
p_hat <- num_edges_obs / choose(population, 2)

fit <- CCM(
  null_model = G_obs ~ Edges,
  alt_model  = ~ DegreeDist,
  population = population,
  Prob_Distr = list("NP"),
  Prob_Distr_Params = list(list(
    dbinom(0:choose(population,2), size = choose(population,2), prob = p_hat)
  )),
  covPattern = rep(0L, population),
  samplesize = 1000L,
  burnin = 200000L,
  interval = 1000L
)

fit
```

```{r plot-degrees}

plot(fit, type = "test")
plot(fit, type = "network")

```

> **Interpretation:** Minimal deviation in degrees indicates that modeling edges alone is sufficient for ER networks.

## Example 2.2: BA Network

```{r structural-ba}
population <- 100
m_attach <- 3
G_obs <- barabasi.game(n = population, m = m_attach, directed = FALSE)

num_edges_obs <- gsize(G_obs)
p_hat <- num_edges_obs / choose(population, 2)

fit <- CCM(
  null_model = G_obs ~ Edges,
  alt_model  = ~ DegreeDist,
  population = population,
  Prob_Distr = list("NP"),
  Prob_Distr_Params = list(list(
    dbinom(0:choose(population,2), size = choose(population,2), prob = p_hat)
  )),
  covPattern = rep(0L, population),
  samplesize = 1000L,
  burnin = 200000L,
  interval = 1000L
)

fit
```


```{r plot-degrees_BA}

plot(fit, type = "test")
plot(fit, type = "network")

```

> **Interpretation:** Large deviations in degree distribution reveal that edges-only Null model is insufficient, highlighting preferential attachment as a structural property.
