---
title: "Assessing Degree Distribution in CCM Networks"
author: "Ravi Goyal"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4
)

# Load libraries
library(CCMnetpy)
library(tidyverse)
library(igraph)
library(reticulate)

# -----------------------------
# Python setup for CCMnet
# -----------------------------

# Check if the environment exists
env_name <- "r-ccmnet" #"ccmnet_env"

if (!env_name %in% reticulate::virtualenv_list() &&
    !env_name %in% reticulate::conda_list()$name) {
  message("CCMnet Python environment not found. Please run install_ccmnet() first. Here is an example of the command: install_ccmnet(envname = env_name, python_version = NULL, restart_session = FALSE)")
}

# Use the environment
reticulate::use_virtualenv(env_name, required = TRUE)

# Initialize CCMnet Python backend
CCMnet_python_setup()
```

## Introduction

This vignette demonstrates how to assess whether the **degree distribution** is part of the generative process of a network using CCMnet. We simulate an observed network, fit a CCM using the observed edge count, and perform a posterior predictive check on the degree distribution.

## 1. Generate Observed Network

```{r observed-network}

# Parameters
population <- 100

# Erdős-Rényi network
p_edge <- 0.05
G_obs <- erdos.renyi.game(n = population, p.or.m = p_edge, type = "gnp", directed = FALSE)

# Degree distribution of observed network
degree_obs <- degree(G_obs)
degree_counts_obs <- table(factor(degree_obs, levels = 0:(population-1)))

# Number of edges and empirical probability
num_edges_obs <- gsize(G_obs)
p_hat <- num_edges_obs / choose(population, 2)

```

## 2. Fit CCM to Observed Network

```{r fit-ccm}
# Fit CCM with Binomial edge distribution

fit <- CCM(
  null_model = G_obs ~ edges,
  alt_model  = ~ degree,
  population = population,
  Prob_Distr = list("NP"),
  Prob_Distr_Params = list(
    dbinom(0:choose(population,2),
           size = choose(population,2),
           prob = p_hat)
  ),
  covPattern = rep(0L, population),
  samplesize = 1000L,
  burnin = 200000L,
  interval = 1000L
)

fit
```

## 3. Visualization

```{r plot-degrees}

plot(fit, type = "test")
plot(fit, type = "network")

```

## 4. Interpretation

A small posterior predictive p-value indicates that the observed degree distribution is unlikely under the CCM, suggesting that **degree distribution is part of the network generative process**. Conversely, a large p-value indicates consistency with the CCM assumptions.

## 1. Generate Observed Network

```{r observed-network_BA}

# Parameters
population <- 100

# Barabási–Albert network
m_attach <- 3  # number of edges per new node
G_obs <- barabasi.game(
  n = population,
  m = m_attach,
  directed = FALSE
)

# Degree distribution of observed network
degree_obs <- degree(G_obs)
degree_counts_obs <- table(factor(degree_obs, levels = 0:(population-1)))

# Number of edges and empirical probability
num_edges_obs <- gsize(G_obs)
p_hat <- num_edges_obs / choose(population, 2)

```

## 2. Fit CCM to Observed Network

```{r fit-ccm_BA}
# Fit CCM with Binomial edge distribution

fit <- CCM(
  null_model = G_obs ~ edges,
  alt_model  = ~ degree,
  population = population,
  Prob_Distr = list("NP"),
  Prob_Distr_Params = list(
    dbinom(0:choose(population,2),
           size = choose(population,2),
           prob = p_hat)
  ),
  covPattern = rep(0L, population),
  samplesize = 1000L,
  burnin = 200000L,
  interval = 1000L
)

fit
```

## 3. Visualization

```{r plot-degrees_BA}

plot(fit, type = "test")
plot(fit, type = "network")

```
