---
title: "CCMnet: Edge, Degree, Mixing, and Structural Assessment"
author: "Ravi Goyal"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCMnet Combined Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4
)

# Load libraries
library("ergm")
library(CCMnet)
library(tidyverse)
library(igraph)
library(gtools)
library(mvtnorm)
library(intergraph)
library(kableExtra)
library(RBesT)
```

# Introduction

```{r data-school}

data("faux.mesa.high")
data("faux.desert.high")
data("faux.dixon.high")
data("faux.magnolia.high")

mesa_net = asIgraph(faux.mesa.high)
desert_net = asIgraph(faux.desert.high)
dixon_net = asIgraph(faux.dixon.high)
magnolia_net = asIgraph(faux.magnolia.high)

# Create summary table
hs_summary <- data.frame(
  High_School = c("Mesa High", "Desert High", "Dixon High", "Magnolia High"),
  Nodes = c(
    vcount(mesa_net),
    vcount(desert_net),
    vcount(dixon_net),
    vcount(magnolia_net)
  ),
  Edges = c(
    gsize(mesa_net),
    gsize(desert_net),
    gsize(dixon_net),
    gsize(magnolia_net)
  )
)

# Render table
kable(
  hs_summary,
  caption = "Summary of high school friendship networks from the ERGM package"
)

```

# Posterior Predictive for New School

``````{r new-school-estimation}

#Beta
prior.unif <- mixbeta(c(1, 1, 1))
r <- sum(hs_summary$Edges); n <- sum(choose(hs_summary$Nodes,2))
posterior.sum_beta <- postmix(prior.unif, n=n, r=r)
print(posterior.sum_beta)

#Normal
sigma <- sd(hs_summary$Edges/choose(hs_summary$Nodes,2))
prior.mean <- .5
prior.se <- 1
prior <- mixnorm(c(1, prior.mean, prior.se), sigma = sigma)
mn <- mean(hs_summary$Edges/choose(hs_summary$Nodes,2))
n <- length(hs_summary$Nodes)
posterior.sum_normal <- postmix(prior, m = mn, n = n)
print(posterior.sum_normal)

```

```{r density-example-new_school-beta}
population <- 500
covPattern <- rep(0L, population)

fit <- CCM_fit(
  Network_stats = list("Density"),
  Prob_Distr = list("Beta"),
  Prob_Distr_Params = list(list(posterior.sum_beta[2], posterior.sum_beta[3])), 
  population = population,
  covPattern = covPattern,
  remove_var_last_entry = FALSE,
  samplesize = 50000L,
  burnin = 200000L,
  interval = 1000L,
)

# Compare MCMC samples to theoretical Poisson distribution
fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))

# Plot density with theoretical distribution
plot(fit, stats = "density", type = "density", include_theoretical = TRUE)
```

```{r density-example-new_school-normal}
population <- 100
covPattern <- rep(0L, population)

fit <- CCM_fit(
  Network_stats = list("Density"),
  Prob_Distr = list("Normal"),
  Prob_Distr_Params = list(list(posterior.sum_normal[2], (posterior.sum_normal[3])^2)), 
  population = population,
  covPattern = covPattern,
  remove_var_last_entry = FALSE,
  samplesize = 10000L,
  burnin = 100000L,
  interval = 1000L,
)

# Compare MCMC samples to theoretical Poisson distribution
fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))

# Plot density with theoretical distribution
plot(fit, stats = "density", type = "density", include_theoretical = TRUE)
```

```{r density-example-new_school-normal-compare}

n_samples <- nrow(fit$mcmc_stats)

net <- network(population, directed = FALSE)
theta = log(posterior.sum_normal[2] / (1-posterior.sum_normal[2]))

ERGM <- simulate(
  net ~ edges,
  coef = theta,
  nsim = n_samples,
  output = "stats"
) / choose(population, 2)

ER = rep(posterior.sum_normal[2], n_samples)

Network_samples <- data.frame(
  value = c(fit$mcmc_stats$density, ERGM, ER),
  model = c(rep("CCMnet", n_samples), rep("ERGM", n_samples), rep("ER", n_samples))
)

ggplot(
  Network_samples %>% filter(model != "ER"),
  aes(x = value, fill = model)
) +
  geom_density(alpha = 0.25, linewidth = .1) +

  geom_vline(
    aes(xintercept = posterior.sum_normal[2], colour = "ER"),
    linewidth = 1,
    linetype = "dotted"
  ) +

  scale_fill_manual(
    values = c(
      CCMnet = "red",
      ERGM   = "green"
    )
  ) +
  scale_colour_manual(
    values = c(
      ER = "blue"
    )
  ) +

  labs(
    x = "Density",
    y = "Probability density",
    fill = "Model",
    colour = NULL   # 🔑 removes duplicate "Model"
  ) +

  guides(
    colour = guide_legend(override.aes = list(fill = NA)),
    fill   = guide_legend(override.aes = list(colour = NA))
  ) +

  theme_bw() +
  theme(legend.position = "top")

```


# Posterior Predictive for Sampled School

```{r density-example-sampled_school-beta}

dixon_deg = degree(dixon_net)
N = length(dixon_deg)

res = data.frame(model = NULL,
                 value = NULL)

for (num_sample_nodes in seq(0,240,40)) {
  dixon_deg_sample = sample(x = dixon_deg, size = num_sample_nodes, replace = FALSE)
  
  #Beta
  prior.unif <- mixbeta(c(1, 1, 1))
  r <- sum(dixon_deg_sample); n <- sum(num_sample_nodes*vcount(dixon_net))
  posterior.sum_beta <- postmix(prior.unif, n=n, r=r)
  
  # Original posterior
  alpha_post <- posterior.sum_beta[2]
  beta_post  <- posterior.sum_beta[3]

  mu <- alpha_post / (alpha_post + beta_post)

  # Infinite-population variance
  var_inf <- (alpha_post * beta_post) /
            ((alpha_post + beta_post)^2 * (alpha_post + beta_post + 1))

  # Finite population correction
  fpc <- (N - num_sample_nodes) / (N - 1)
  var_fpc <- var_inf * fpc

  # Moment-matched Beta parameters
  S <- mu * (1 - mu) / var_fpc - 1
  posterior.sum_beta[2] <- mu * S
  posterior.sum_beta[3] <- (1 - mu) * S
  
  population <- vcount(dixon_net)
  covPattern <- rep(0L, population)
  
  fit <- CCM_fit(
    Network_stats = list("Density"),
    Prob_Distr = list("Beta"),
    Prob_Distr_Params = list(list(posterior.sum_beta[2], posterior.sum_beta[3])), 
    population = population,
    covPattern = covPattern,
    remove_var_last_entry = FALSE,
    samplesize = 10000L,
    burnin = 100000L,
    interval = 1000L,
  )
  res = bind_rows(res, data.frame(
    model = rep("CCMnet", length(fit$mcmc_stats$density)),
    value = fit$mcmc_stats$density,
    sample = num_sample_nodes))
  
  n_samples <- nrow(fit$mcmc_stats)
  
  net <- network(population, directed = FALSE)
  p = posterior.sum_beta[2]/(posterior.sum_beta[2] + posterior.sum_beta[3])
  theta = log(p / (1-p))

  ERGM <- simulate(
    net ~ edges,
    coef = theta,
    nsim = n_samples,
    output = "stats"
  ) / choose(population, 2)

  ER = rep(posterior.sum_beta[2]/(posterior.sum_beta[2] + posterior.sum_beta[3]), n_samples)
  
  res <- bind_rows(res, data.frame(
    value = c(ERGM, ER),
    model = c(rep("ERGM", n_samples), rep("ER", n_samples)),
    sample = c(rep(num_sample_nodes, n_samples), rep(num_sample_nodes, n_samples))
  ))
  
}


```

```{r density-example-sampled_school-compare}

# ER vertical lines
ER_lines <- res %>%
  filter(model == "ER") %>%
  group_by(sample, model) %>%
  summarise(xintercept = unique(value), .groups = "drop")

ggplot(
  res %>% filter(model != "ER"),
  aes(x = value, fill = model)
) +
  geom_density(alpha = 0.25, linewidth = .1) +

  # ER vertical lines
  geom_vline(
    data = ER_lines,
    aes(xintercept = xintercept, colour = model),
    linetype = "dotted",
    linewidth = 1
  ) +

  scale_fill_manual(
    values = c(
      CCMnet = "red",
      ERGM   = "green"
    )
  ) +
  scale_colour_manual(
    values = c(
      ER = "blue"
    )
  ) +

  labs(
    x = "Density",
    y = "Probability density",
    fill = "Model",
    colour = NULL
  ) +

  guides(
    colour = guide_legend(override.aes = list(fill = NA)),
    fill   = guide_legend(override.aes = list(colour = NA))
  ) +

  theme_bw() +
  theme(legend.position = "top") +
  facet_wrap(~sample, scales =  "free")

```