---
title: "CCMnet: Edge, Degree, Degree Mixing, and Triangles"
author: "Ravi Goyal"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCMnet Combined Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4
)

# Load libraries
library(CCMnet)
library(tidyverse)
library(igraph)
library(gtools)
library(mvtnorm)
```

# Introduction

This vignette demonstrates the use of **Congruence Class Models (CCMs)** for network modeling. CCMs allow us to define a generative process for networks in terms of chosen network properties (edges, degree distributions, mixing matrices) and generate networks consistent with these properties.

The main purpose is to:

**Verification of the MCMC sampler**: We sample networks according to known distributions for edges, degrees, and mixing matrices, and verify that the CCM sampler reproduces these distributions.


# Section 1: Edge Count Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific edge distributions.

## Example 1.1: Poisson Edge Count

We fit a CCM where the **number of edges** follows a Poisson distribution with \(\lambda = 350\). The probability of sampling a network with k edges is given by the Poisson mass function:

\[ P(K=k) = \frac{\lambda^k e^{-\lambda}}{k!} \]

```{r edge-example-poisson}

ccm_sample <- sample_ccm(
  network_stats = list("edges"),
  prob_distr = list("poisson"),
  prob_distr_params = list(list(350)), 
  population = 100
)

# Compare MCMC samples to theoretical Poisson distribution
ccm_sample<- CCM_theoretical_check(ccm_sample, n_sim = nrow(ccm_sample$mcmc_stats))

#summary
summary(ccm_sample)

#print
print(ccm_sample)

# Plot density with theoretical distribution
plot(ccm_sample, stats = "edges", type = "hist", include_theoretical = TRUE)

#TEMP
#p <- plot(fit, stats = "edges", type = "density", include_theoretical = TRUE)
#ggsave("/Users/ravigoyal/Dropbox/Academic/Research/Projects/CCMnet_package_manuscript/Figures/edges_pois.pdf", plot = p, width = 7, height = 7, units = "cm")
```

> **Explanation:** The density plot shows the CCM MCMC samples (blue) and the theoretical Poisson distribution (red). Close overlap verifies correct sampling.

## Example 1.2: Uniform Edge Count

```{r edge-example-uniform}

ccm_sample <- sample_ccm(
  network_stats = list("edges"),
  prob_distr = list("uniform"), 
  prob_distr_params = list(list(0)),
  population = 50,
  sample_size = 50000L,
  burnin = 200000L,
  interval = 1000L,
)

ccm_sample<- CCM_theoretical_check(ccm_sample, n_sim = nrow(ccm_sample$mcmc_stats))
summary(ccm_sample)
print(ccm_sample)
plot(ccm_sample, stats = "edges", type = "hist", include_theoretical = TRUE)

#TEMP
#p <- plot(fit, stats = "edges", type = "density", include_theoretical = TRUE)
#ggsave("/Users/ravigoyal/Dropbox/Academic/Research/Projects/CCMnet_package_manuscript/Figures/edges_unif.pdf", plot = p, width = 7, height = 7, units = "cm")
```

> **Explanation:** The density plot verifies that CCM samples uniformly reproduce all possible edge counts.

## Example 1.3: Non-Parametric Edge Count

```{r edge-example-np}

population = 100

prob_distr_params_P1 <- dpois(0:choose(population,2), lambda = 50)
prob_distr_params_P2 <- dpois(0:choose(population,2), lambda = 100)
prob_distr_params_C <- prob_distr_params_P1 + prob_distr_params_P2
prob_distr_params <- prob_distr_params_C / sum(prob_distr_params_C)

ccm_sample <- sample_ccm(
  network_stats = list("edges"),
  prob_distr = list("np"),
  prob_distr_params = list(list(prob_distr_params)),
  population = 100,
  sample_size = 50000L,
  burnin = 200000L
)

ccm_sample<- CCM_theoretical_check(ccm_sample, n_sim = nrow(ccm_sample$mcmc_stats))
summary(ccm_sample)
print(ccm_sample)
plot(ccm_sample, stats = "edges", type = "hist", include_theoretical = TRUE)

#TEMP
#p <- plot(fit, stats = "edges", type = "density", include_theoretical = TRUE)
#ggsave("/Users/ravigoyal/Dropbox/Academic/Research/Projects/CCMnet_package_manuscript/Figures/edges_np.pdf", plot = p, width = 7, height = 7, units = "cm")
```

# Section 2: Degree Distribution Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific degree distribution.

```{r degree-example-dirmult}

# Network_stats ='DegreeDist'
# Prob_Distr ='DirMult'
# Prob_Distr_Params =list(list(c(2,21,15,12)))
# samplesize  = 10000
# burnin =100000
# interval =1000
# statsonly = TRUE
# G = NULL
# population = 500
# covPattern = NULL
# remove_var_last_entry = FALSE
# Obs_stats = NULL

ccm_sample<- sample_ccm(network_stats='DegreeDist',
               prob_distr='DirMult',
               prob_distr_params=list(list(c(2,21,15,12))), 
               population = 500, 
               sample_size = 10000,
               burnin=100000, 
               interval=1000) 

ccm_sample <- CCM_theoretical_check(ccm_sample, n_sim = 100)
summary(ccm_sample)
print(ccm_sample)
plot(ccm_sample, stats = paste0("deg", 0:3), type = "hist", include_theoretical = TRUE)
    
#TEMP
#p <- plot(fit, stats = paste0("deg", 0:3), type = "density", include_theoretical = TRUE)
#ggsave("/Users/ravigoyal/Dropbox/Academic/Research/Projects/CCMnet_package_manuscript/Figures/degreedist_dirmult.pdf", plot = p)
```

> **Explanation:** The density plot verifies that CCM samples reproduce the correct degre distribution.

# Section 3: Degree Mixing Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific degree mixing distribution.

```{r degmix-example-normal}
population <- 500
covPattern <- rep(0L, population)

mean_vec = c(23.05, 66.05, 44.05, 91.05, 118.05, 80.05)

var_mat = matrix(data = c(22.865223, -3.602583, -2.402631, -4.966165, -6.438834, -4.366189,
                          -3.602583, 58.463833, -6.884763, -14.230595, -18.450541, -12.511358,
                          -2.402631, -6.884763, 41.398484, -9.490654, -12.305016, -8.344062,
                          -4.966165, -14.230595, -9.490654, 74.936848, -25.434092, -17.246921,
                          -6.438834, -18.450541, -12.305016, -25.434092, 89.239302, -22.361329,
                          -4.366189, -12.511358, -8.344062, -17.246921, -22.361329, 68.071353), ncol = 6)
Prob_Distr_Params = list(list(mean_vec, var_mat))

ccm_sample <- sample_ccm(network_stats='DegMixing',
                              prob_distr='Normal',
                              prob_distr_params=Prob_Distr_Params, 
                              sample_size = 50000,
                              burnin=100000, 
                              interval=1000,
                              population=500, 
                              cov_pattern = covPattern) 

ccm_sample <- CCM_theoretical_check(ccm_sample, n_sim = 100)
summary(ccm_sample)
print(ccm_sample)
plot(ccm_sample, stats = c("DM11", "DM12", "DM13", "DM22", "DM23", "DM33"), type = "density", include_theoretical = TRUE)

```

# Section 4: Degree Mixing and Triangle Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific degree mixing and triangle distribution.

```{r degmix_triangle-example-normal}
population <- 500
covPattern <- rep(0L, population)

mean_vec = c(23.05, 66.05, 44.05, 91.05, 118.05, 80.05)

var_mat = matrix(data = c(22.865223, -3.602583, -2.402631, -4.966165, -6.438834, -4.366189,
                          -3.602583, 58.463833, -6.884763, -14.230595, -18.450541, -12.511358,
                          -2.402631, -6.884763, 41.398484, -9.490654, -12.305016, -8.344062,
                          -4.966165, -14.230595, -9.490654, 74.936848, -25.434092, -17.246921,
                          -6.438834, -18.450541, -12.305016, -25.434092, 89.239302, -22.361329,
                          -4.366189, -12.511358, -8.344062, -17.246921, -22.361329, 68.071353), ncol = 6)
Prob_Distr_Params = list(list(mean_vec, var_mat),
                         list(6,3))

# Network_stats =c('DegMixing', 'Triangles')
# Prob_Distr =c('Normal', 'Normal')
# Prob_Distr_Params =Prob_Distr_Params
# samplesize  = 100000
# burnin =1000000
# interval =1000
# statsonly = TRUE
# G = NULL
# population = 500
# covPattern = NULL
# remove_var_last_entry = FALSE
# Obs_stats = NULL

ccm_sample <- sample_ccm(network_stats=c('DegMixing', 'Triangles'),
                              prob_distr=c('Normal', 'Normal'),
                              prob_distr_params=Prob_Distr_Params, 
                              sample_size = 100000,
                              burnin=1000000, 
                              interval=1000,
                              population=500, 
                              cov_pattern = covPattern) 

ccm_sample <- CCM_theoretical_check(ccm_sample, n_sim = 100)
summary(ccm_sample)
print(ccm_sample)
CCM_traceplot(ccm_sample, stats = "triangles")
plot(ccm_sample, stats = c("DM11", "DM12", "DM13", "DM22", "DM23", "DM33", "triangles"), type = "hist", include_theoretical = TRUE)

#TEMP
#p <- plot(fit, stats = c("DM11", "DM12", "DM13", "DM22", "DM23", "DM33", "triangles"), type = "density", include_theoretical = TRUE)
#ggsave("/Users/ravigoyal/Dropbox/Academic/Research/Projects/CCMnet_package_manuscript/Figures/degmix_triangles_normal.pdf", plot = p)

#p <- CCM_traceplot(fit, stats = "triangles")
#ggsave("/Users/ravigoyal/Dropbox/Academic/Research/Projects/CCMnet_package_manuscript/Figures/degmix_triangles_normal_traceplot.pdf", plot = p)
```

> **Explanation:** The density plots show sampled degrees (blue) versus theoretical probabilities (red). Close match verifies correct MCMC sampling.
