---
title: "CCMnet: Mixing and Degree Distribution+Mixing"
author: "Ravi Goyal"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCMnet Combined Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4
)

# Load libraries
library(CCMnet)
library(tidyverse)
library(igraph)
library(gtools)
library(mvtnorm)
```

# Introduction

This vignette demonstrates the use of **Congruence Class Models (CCMs)** for network modeling. CCMs allow us to define a generative process for networks in terms of chosen network properties (edges, degree distributions, mixing matrices) and generate networks consistent with these properties.

The main purpose is to:

**Verification of the MCMC sampler**: We sample networks according to known distributions for edges, degrees, and mixing matrices, and verify that the CCM sampler reproduces these distributions.


# Section 1: Characteristic Mixing Matrix Sampler Verification

```{r mixing-example-poisson}
population <- 100
covPattern <- c(rep(1L, population/2), rep(2L, population/2))

Network_stats = list("Mixing")
Prob_Distr = list('Poisson')
Prob_Distr_Params=list(list(c(21,15,12)))
population = population
covPattern = covPattern
remove_var_last_entry = FALSE

samplesize = 1000L
burnin = 1000L
interval = 1000L
statsonly = TRUE
G = NULL
Obs_stats = NULL

fit <- CCM_fit(
  Network_stats = list("Mixing"),
  Prob_Distr = list('Poisson'),
  Prob_Distr_Params=list(list(c(21,15,12))), 
  population = population,
  covPattern = covPattern,
  remove_var_last_entry = FALSE
)

# Compare MCMC samples to theoretical Poisson distribution
fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))

# Plot density with theoretical distribution
plot(fit, stats = c("M11", "M21", "M22"), type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plot shows the CCM MCMC samples (blue) and the theoretical Poisson distribution (red). Close overlap verifies correct sampling.

## Example 2: Uniform Edge Count

```{r mixing-degdist-example-uniform}

population <- 100
covPattern <- c(rep(1L, population/2), rep(2L, population/2))

var_mat = diag(rep(2,4))
Prob_Distr_Params=list(list(list(c(5,10,25,10), c(10,5,20,15)), list(var_mat, var_mat)), 
                       list(c(30), c(2)))

fit <- CCM_fit(
  Network_stats = c("DegreeDist", "Mixing"),
  Prob_Distr =  c("Normal", "Normal"), 
  Prob_Distr_Params = Prob_Distr_Params,
  population = 100,
  covPattern =  covPattern,
  remove_var_last_entry = FALSE,
  samplesize = 50000L,
  burnin = 100000L,
  interval = 1000L,
)

#Diagnostics
CCM_traceplot(fit, stats = "M21")

# Compare MCMC samples to theoretical Poisson distribution
fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))

# Plot density with theoretical distribution

stat_list = c(paste(paste0("deg", 0:(3)), "_1", sep = ""),
              paste(paste0("deg", 0:(3)), "_2", sep = ""),
              c("M21"))
    
plot(fit, stats = stat_list, type = "density", include_theoretical = TRUE)
```

