---
title: "CCMnet: Edge, Degree, Mixing, and Structural Assessment"
author: "Ravi Goyal"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCMnet Combined Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4
)

# Load libraries
library(CCMnetpy)
library(tidyverse)
library(igraph)
library(reticulate)

# -----------------------------
# Python setup for CCMnet
# -----------------------------

env_name <- "r-ccmnet"

if (!env_name %in% reticulate::virtualenv_list() &&
    !env_name %in% reticulate::conda_list()$name) {
  message("CCMnet Python environment not found. Please run install_ccmnet() first. Here is an example of the command: install_ccmnet(envname = env_name, python_version = NULL, restart_session = FALSE)")
}

reticulate::use_virtualenv(env_name, required = TRUE)

CCMnet_python_setup()
```

# Introduction

This vignette demonstrates the use of **Congruence Class Models (CCMs)** for network modeling. CCMs allow us to define a generative process for networks in terms of chosen network properties (edges, degree distributions, mixing matrices) and generate networks consistent with these properties.

We distinguish two main purposes:

1. **Verification of the MCMC sampler** (Sections 1–3): We sample networks according to known distributions for edges, degrees, and mixing matrices, and verify that the CCM sampler reproduces these distributions.
2. **Structural assessment** (Section 4): We test whether a network property (e.g., degree distribution) is part of the generative process by comparing observed networks to Null CCMs.


# Section 1: Edge Count Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific edge distributions.

## Example 1.1: Poisson Edge Count

We fit a CCM where the **number of edges** follows a Poisson distribution with \(\lambda = 6\). The probability of sampling a network with k edges is given by the Poisson mass function:

\[ P(K=k) = \frac{\lambda^k e^{-\lambda}}{k!} \]

```{r edge-example-poisson}
population <- 100
covPattern <- rep(0L, population)

fit <- CCM_fit(
  Network_stats = list("Edge"),
  Prob_Distr = list("Poisson"),
  Prob_Distr_Params = list(6),
  population = population,
  covPattern = covPattern
)

# Compare MCMC samples to theoretical Poisson distribution
fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))

# Plot density with theoretical distribution
plot(fit, stats = "edges", type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plot shows the CCM MCMC samples (blue) and the theoretical Poisson distribution (red). Close overlap verifies correct sampling.

## Example 1.2: Uniform Edge Count

```{r edge-example-uniform}
population <- 10
covPattern <- rep(0L, population)

fit <- CCM_fit(
  Network_stats = list("Edge"),
  Prob_Distr = list("uniform"),
  Prob_Distr_Params = NULL,
  population = population,
  covPattern = covPattern
)

fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))
plot(fit, stats = "edges", type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plot verifies that CCM samples uniformly reproduce all possible edge counts.

## Example 1.3: Non-Parametric Edge Count

```{r edge-example-np}
population <- 100
covPattern <- rep(0L, population)

Prob_Distr_Params_P1 <- dpois(0:choose(population,2), lambda = 5)
Prob_Distr_Params_P2 <- dpois(0:choose(population,2), lambda = 10)
Prob_Distr_Params_C <- Prob_Distr_Params_P1 + Prob_Distr_Params_P2
Prob_Distr_Params_C <- Prob_Distr_Params_C / sum(Prob_Distr_Params_C)

fit <- CCM_fit(
  Network_stats = list("Edge"),
  Prob_Distr = list("NP"),
  Prob_Distr_Params = list(Prob_Distr_Params_C),
  population = population,
  covPattern = covPattern,
  samplesize = 10000L
)

fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))
plot(fit, stats = "edges", type = "density", include_theoretical = TRUE)
```

> **Explanation:** Non-parametric sampling reproduces the combined Poisson distribution, verifying the CCM sampler works for arbitrary edge distributions.


# Section 2: Degree Distribution Sampler Verification

Next, we verify that the CCM sampler reproduces specified degree distributions using multinomial or multinomial-Poisson models.

## Example 2.1: Multinomial-Poisson Degree Distribution

```{r degree-example}
population <- 100
covPattern <- rep(0L, population)

Network_stats <- list("Degree")
Prob_Distr <- list("Multinomial_Poisson")
Prob_Distr_Params <- list(
  population,
  dnbinom(0:(population-1), size = 1.017340, mu = 6.192894)
)

fit <- CCM_fit(
  Network_stats = Network_stats,
  Prob_Distr = Prob_Distr,
  Prob_Distr_Params = Prob_Distr_Params,
  population = as.integer(population),
  covPattern = as.integer(covPattern)
)

fit <- CCM_theoretical_check(fit)
plot(fit, stats = paste0("deg", 0:9), type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plots show sampled degrees (blue) versus theoretical multinomial-Poisson probabilities (red). Close match verifies correct MCMC sampling.


# Section 3: Mixing Matrix Sampler Verification

We verify that CCM samples reproduce a specified mixing matrix.

```{r mixing-example}
population <- 100
covPattern <- c(rep(0L, population/2), rep(1L, population/2))

Network_stats <- list("Mixing")
Prob_Distr <- list("Multinomial_Poisson")
Prob_Distr_Params <- list(
  20,           # mean number of mixing edges
  c(.3, .6, .1) # probabilities for M11, M21, M22
)

fit <- CCM_fit(
  Network_stats = Network_stats,
  Prob_Distr = Prob_Distr,
  Prob_Distr_Params = Prob_Distr_Params,
  population = population,
  covPattern = covPattern
)

fit <- CCM_theoretical_check(fit)
plot(fit, stats = c("M11","M21","M22"), type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plots verify that the MCMC sampler reproduces the intended mixing probabilities.


# Section 4: Structural Assessment

Structural assessment tests whether a network property is part of the generative process.

- **Erdős-Rényi (ER) Network:** Only the number of edges is relevant; degree distribution deviations should be minimal.
- **Barabási–Albert (BA) Network:** Preferential attachment induces heterogeneous degrees; a Null model with edges only cannot capture this.

## Example 4.1: ER Network

```{r structural-er}
population <- 100
p_edge <- 0.05
G_obs <- erdos.renyi.game(n = population, p.or.m = p_edge, type = "gnp", directed = FALSE)

# Number of edges and empirical probability
num_edges_obs <- gsize(G_obs)
p_hat <- num_edges_obs / choose(population, 2)

fit <- CCM(
  null_model = G_obs ~ edges,
  alt_model  = ~ degree,
  population = population,
  Prob_Distr = list("NP"),
  Prob_Distr_Params = list(
    dbinom(0:choose(population,2), size = choose(population,2), prob = p_hat)
  ),
  covPattern = rep(0L, population),
  samplesize = 1000L,
  burnin = 200000L,
  interval = 1000L
)

fit
```

```{r plot-degrees}

plot(fit, type = "test")
plot(fit, type = "network")

```

> **Interpretation:** Minimal deviation in degrees indicates that modeling edges alone is sufficient for ER networks.

## Example 4.2: BA Network

```{r structural-ba}
population <- 100
m_attach <- 3
G_obs <- barabasi.game(n = population, m = m_attach, directed = FALSE)

num_edges_obs <- gsize(G_obs)
p_hat <- num_edges_obs / choose(population, 2)

fit <- CCM(
  null_model = G_obs ~ edges,
  alt_model  = ~ degree,
  population = population,
  Prob_Distr = list("NP"),
  Prob_Distr_Params = list(
    dbinom(0:choose(population,2), size = choose(population,2), prob = p_hat)
  ),
  covPattern = rep(0L, population),
  samplesize = 1000L,
  burnin = 200000L,
  interval = 1000L
)

fit
```


```{r plot-degrees_BA}

plot(fit, type = "test")
plot(fit, type = "network")

```

> **Interpretation:** Large deviations in degree distribution reveal that edges-only Null model is insufficient, highlighting preferential attachment as a structural property.
