---
title: "CCMnet: Edge, Degree, Mixing, and Structural Assessment"
author: "Ravi Goyal"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCMnet Combined Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4
)

# Load libraries
library(CCMnet)
library(tidyverse)
library(igraph)
library(gtools)
library(mvtnorm)
```

# Introduction

This vignette demonstrates the use of **Congruence Class Models (CCMs)** for network modeling. CCMs allow us to define a generative process for networks in terms of chosen network properties (edges, degree distributions, mixing matrices) and generate networks consistent with these properties.

We distinguish two main purposes:

1. **Verification of the MCMC sampler** (Sections 1–4): We sample networks according to known distributions for edges, degrees, and mixing matrices, and verify that the CCM sampler reproduces these distributions.
2. **Structural assessment** (Section 5): We test whether a network property (e.g., degree distribution) is part of the generative process by comparing observed networks to Null CCMs.


# Section 1: Edge Count Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific edge distribution.

```{r edge-example-normal}
population <- 100
covPattern <- rep(0L, population)

Network_stats = list("Density")
Prob_Distr = list("Normal")
Prob_Distr_Params = list(list(.04, .0001))
population = 100
samplesize = 1000L
burnin = 200000L
interval = 1000L
covPattern = rep(0L, population)
statsonly = TRUE
G = NULL
P = NULL
use_G = FALSE
outfile = "none"
partial_network = as.integer(0)
obs_nodes = NULL
MH_proposal_type = "TNT"
Obs_stats = NULL
remove_var_last_entry = FALSE
bayesian_inference = FALSE
Ia = NULL
Il = NULL
R = NULL
epi_params = NULL
print_calculations = FALSE

fit <- CCM_fit(
  Network_stats = list("Density"),
  Prob_Distr = list("Normal"),
  Prob_Distr_Params = list(list(.04, .0001)),
  population = population,
  covPattern = covPattern,
  remove_var_last_entry = FALSE
)

# Compare MCMC samples to theoretical Poisson distribution
fit <- CCM_theoretical_check(fit, n_sim = nrow(fit$mcmc_stats))

# Plot density with theoretical distribution
plot(fit, stats = "density", type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plot shows the CCM MCMC samples (blue) and the theoretical Poisson distribution (red). Close overlap verifies correct sampling.

# Section 2: Degree Distribution Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific degree distribution.

```{r degree-example-dirmult}
population <- 500
covPattern <- rep(0L, population)

fit <- CCM_fit(Network_stats='DegreeDist',
               Prob_Distr='DirMult',
               Prob_Distr_Params=list(list(c(2,21,15,12))), 
               population=population, 
               covPattern = covPattern,
               samplesize = 10000,
               burnin=100000, 
               interval=1000,
               remove_var_last_entry = FALSE) 

fit$mcmc_stats %>% apply(2, mean)

fit <- CCM_theoretical_check(fit, n_sim = 100)
plot(fit, stats = paste0("deg", 0:3), type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plot verifies that CCM samples reproduce the correct degre distribution.

# Section 3: Degree Mixing Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific degree mixing distribution.

```{r degmix-example-normal}
population <- 500
covPattern <- rep(0L, population)

mean_vec = c(23.05, 66.05, 44.05, 91.05, 118.05, 80.05)

var_mat = matrix(data = c(22.865223, -3.602583, -2.402631, -4.966165, -6.438834, -4.366189,
                          -3.602583, 58.463833, -6.884763, -14.230595, -18.450541, -12.511358,
                          -2.402631, -6.884763, 41.398484, -9.490654, -12.305016, -8.344062,
                          -4.966165, -14.230595, -9.490654, 74.936848, -25.434092, -17.246921,
                          -6.438834, -18.450541, -12.305016, -25.434092, 89.239302, -22.361329,
                          -4.366189, -12.511358, -8.344062, -17.246921, -22.361329, 68.071353), ncol = 6)
Prob_Distr_Params = list(list(mean_vec, var_mat))

fit = CCM_fit(Network_stats='DegMixing',
                              Prob_Distr='Normal',
                              Prob_Distr_Params=Prob_Distr_Params, 
                              samplesize = 50000,
                              burnin=100000, 
                              interval=1000,
                              population=500, 
                              covPattern = covPattern,
                              remove_var_last_entry = FALSE) 

fit <- CCM_theoretical_check(fit, n_sim = 100)
plot(fit, stats = c("DM11", "DM12", "DM13", "DM22", "DM23", "DM33"), type = "density", include_theoretical = TRUE)
```

# Section 4: Degree Mixing and Triangle Sampler Verification

Here we verify that the CCM sampler correctly reproduces networks with specific degree mixing and triangle distribution.

```{r degmix_triangle-example-normal}
population <- 500
covPattern <- rep(0L, population)

mean_vec = c(23.05, 66.05, 44.05, 91.05, 118.05, 80.05)

var_mat = matrix(data = c(22.865223, -3.602583, -2.402631, -4.966165, -6.438834, -4.366189,
                          -3.602583, 58.463833, -6.884763, -14.230595, -18.450541, -12.511358,
                          -2.402631, -6.884763, 41.398484, -9.490654, -12.305016, -8.344062,
                          -4.966165, -14.230595, -9.490654, 74.936848, -25.434092, -17.246921,
                          -6.438834, -18.450541, -12.305016, -25.434092, 89.239302, -22.361329,
                          -4.366189, -12.511358, -8.344062, -17.246921, -22.361329, 68.071353), ncol = 6)
Prob_Distr_Params = list(list(mean_vec, var_mat),
                         list(6,2))

fit = CCM_fit(Network_stats=c('DegMixing', 'Triangles'),
                              Prob_Distr=c('Normal', 'Normal'),
                              Prob_Distr_Params=Prob_Distr_Params, 
                              samplesize = 50000,
                              burnin=100000, 
                              interval=1000,
                              population=500, 
                              covPattern = covPattern,
                              remove_var_last_entry = FALSE) 

fit$mcmc_stats %>% apply(2, mean)

fit <- CCM_theoretical_check(fit, n_sim = 100)
plot(fit, stats = c("DM11", "DM12", "DM13", "DM22", "DM23", "DM33", "triangles"), type = "density", include_theoretical = TRUE)
```

> **Explanation:** The density plots show sampled degrees (blue) versus theoretical probabilities (red). Close match verifies correct MCMC sampling.

# Section 5: Structural Assessment

Structural assessment tests whether a network property is part of the generative process.

- **Erdős-Rényi (ER) Network:** Only the number of edges is relevant; degree distribution deviations should be minimal.
- **Barabási–Albert (BA) Network:** Preferential attachment induces heterogeneous degrees; a Null model with edges only cannot capture this.

## Example 5.1: ER Network
